``Interestingness'' in Mathematical Theory Exploration

### 2018-05-17 PhDThesis/related_formalstats.tex
More deduplication

\begin{itemize}
\item \emph{Feature vectors} (ordered sets of features) are chosen to be more
  compact than the data they're extracted from: feature extraction is
  \emph{lossy compression}. This reduces the size of the ML problem, improving
  efficiency (e.g. running time).
\item We avoid learning irrelevant details such as the encoding system used,
  improving \emph{data} efficiency (the number of samples required to spot a
  pattern).
\end{itemize}

@@

\paragraph{Sequencing}

@@

@@

@@

Machine learning over structured data:
1D is common: parsing natural language
2D is common; images
Trees are fractal
Backpropagation through structure
LSTM with recursive structure
Most work tries to identify structure; we already have it

Recurrent neural networks
Backpropagation through structure

>>>>

\begin{itemize}
\item \emph{Feature vectors} (ordered sets of features) are chosen to be more
  compact than the data they're extracted from: feature extraction is
  \emph{lossy compression}. This reduces the size of the ML problem, improving
  efficiency (e.g. running time).
\item We avoid learning irrelevant details such as the encoding system used,
  improving \emph{data} efficiency (the number of samples required to spot a
  pattern).
\item All of our feature vectors will be the same size, i.e. they will all have
  length (or \emph{dimension}) $d$. Many machine learning algorithms only work
  with inputs of a uniform size; feature extraction allows us to use these
  algorithms in domains where the size of each input is not known, may vary or
  may even be unbounded. For example, element-wise comparison of feature vectors
  is trivial (compare the $i$th elements for $1 \leq i \leq d$); for expressions
  this is not so straightforward, as their nesting may give rise to very
  different shapes.
\item Unlike our expressions, which are discrete, we can continuously transform
  one feature vector into another. This enables many powerful machine learning
  algorithms to be used, such as those based on \emph{gradient descent} or, in
  our case, arithmetic means.
\item Feature vectors can be chosen to represent the relevant information in a
  more compressed form than the raw data; for example, we might replace verbose,
  descriptive identifiers with sequential numbers. This reduces the input size
  of the machine learning problem, improving efficiency.
\end{itemize}

@@

\paragraph{Structured Data}

@@

Using sequences to represent recursive structures is also problematic: if we
want our learning algorithm to exploit structure (such as the depth of a token),
it will have to discover how to parse the sequences for itself, which seems
wasteful. The \emph{back-propagation through structure} approach
\cite{goller1996learning} is a more direct solution to this problem, using a
feed-forward NN to learn recursive distributed representations
\cite{pollack1990recursive} which correspond to the recursive structure of the
inputs. Such distributed representations can also be used for sequences, which
we can use to encode sub-trees when the branching factor of nodes is not uniform
\cite{kwasny1995tail}. More recent work has investigated storing recursive
structures inside LSTM cells \cite{zhu2015long}.

A simpler alternative for generating recursive distributed representations is to
use circular convolution \cite{conf/ijcai/Plate91}. Although promising results
are shown for its use in \emph{distributed tree kernels}
\cite{zanzotto2012distributed}, our preliminary experiments in applying
circular convolution to functional programming expressions found most of the
information to be lost in the process; presumably as the expressions are too
small.

@@

\emph{Kernel methods} have also been applied to structured information, for
example in \cite{Gartner2003} the input data (including sequences, trees and
graphs) are represented using \emph{generative models}, such as hidden Markov
models, of a fixed size suitable for learning. Many more applications of kernel
methods to structured domains are given in \cite{bakir2007predicting}, which
could be used to learn more subtle relations between expressions than recurrent
clustering alone.

@@
